### 2013-12-04 paper/paper.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}

@@

\section{Introduction}

@@

Recently, Gao and Scholl presented a set of studies \citep{gao09,gao10,gao11} demonstrating that certain aspects of motion influence the detection of chasing and performance in interactive tasks where subjects controlled chasee with mouse. 
In \citet{gao10} the authors focused on the contribution of agent's orientation. 

@@

For example, in Experiment 2 the subjects controlled a green disc with computer mouse and tried to avoid contact with a white circle which chased the green circle. 

@@

As a manipulation the authors added seven white darts that were in half of the trials oriented towards the green circle and in other half perpendicular to it. 

@@

They compared across the two conditions the proportion of trials where chaser caught the green circle (relative to trials where he didn't). 

@@

In Experiment 4 the chaser was the only circle in the display, so it was easily identified. 
However the subjects needed to additionaly avoid contact with darts. 
Escape rate was lower in trials where the darts were oriented towards green circle than in trials where the orientation was perpendicular. 

@@

The authors named the consistent influence of orientation, the wolfpack effect.\\

@@

This is illustrated in figure \ref{fig.dart} on the example of darts used by \citet{gao10}. 
The authors designated the concave vertex (red) as the nominal point for rotation. 
Intuitively however, the center of mass (blue) looks like a much better choice. 
What happens if the perceived center of rotation is displaced in the direction of center of mass.
In this case the rotation around the nominal center is perceived as motion towards chasee and we can contrive an alternative explanation for the results in \citet{gao10}. 
In Experiment 3a the subjects avoided areas with darts oriented towards the prey because the agents' positions were perceived as shifting closer to the green circle. 
In Experiment 4 if we assume that subjects estimated the critical distance to surrounding agents and based their decisions where to move the green circle on this estimate. 
If the perceived agents' location was shifted towards chasee in wolfpack trials, the subjects would often prematurely leave the current location in exchange for another location which would be a bad choice from the point of view of the nominal dart's center which was used to decide whether the green circle was caught or not (see Figure \ref{fig.exp4}). 

@@

Still, since the the intended orientation change is perceived as motion, the results can be explained by the influence of motion cues on subjects judgment.\\
Experiment 3b in \citet{gao10} was the only experiment that did not use darts. 
The design and results were similar to that of Experiment 3a, except that instead of darts the subject tried to avoid white circles whose orientation was determined by two red dots ('eyes', see figure \ref{fig.eyes}).

@@

If we use the shape of the agent to determine the center of mass, this is identical to the point around which the circle was rotated. 

@@

The literature on memory displacement is relevant here \citep{hubbard05}. 

@@

Crucially, some studies \citep{freyd92}[e.g.] demonstrated that displacement can also be influenced by information about stimulus animacy. 
The memory displacement is studied by asking subjects to locate the last position of an object that just disappeared.
This is usually done by asking subjects to select this position with mouse or alternatively to by selecting which of two probes is nearer to the location where the object disappeared.
How are such displacements in explicit recall relevant to the implicit interactive tasks used in \citet{gao10}.
According to \citet{hubbard05}[p. 844] "displacement occurs because it aids in the spatial localization of physical objects and facilitates rapid motor responding to objects in the environment". 
He further adds (cf.) that "accurate spatial localization is important for calibrating an observer's response to a stimulus so that a maximally effective and adaptive interaction with that stimulus might be achieved". 

@@

We can apply this idea to stimuli and tasks in \citet{gao10}. Since agent's head/nose orientation usually predicts the subsequent motion, orientation may be used to predict future position. 

@@

Hence we expect to find memory displacement which is influenced by factors that provide cue to the future motion direction. 

@@

Due to the perceived displacement the subject would avoid the wolfpack areas because the wolfpack stimuli are perceived as closing in on the green circle.\\
In the current study we investigate the influence of perceived displacement on subject's performance in the interactive tasks from \citet{gao10}. 
In particular we chose to revisit their task from Experiment 3, the so-called leave-me-alone task in which subjects avoided contact with randomly moving objects in a display separated into areas with different orientation. 
We chose this Experiment since it was the only one that demonstrated the wolfpack effect with bug stimuli (displayed in figure \ref{fig.eyes}) where there is no obvious reason for discrepancy between the nominal and perceived object's center (as is the case darts and their shifted center of mass).\\

@@

  \item This displacement influences subjects' performance in the leave-me-alone task. 

@@

To test the second claim we shift the rotation center along the anteroposterior axis of the agent and measure how this manipulation influences subject's avoidance of wolfpack areas. 
We utilize two tasks to test the first claim. 
First, we append a location recall task to the leave-me-alone task. 
After each trial of leave-me-alone task one dart/bug disappears while all the remaining objects freezee at their last position. 
The subject is asked to select the position where the agent disappeared.
Second, we devised a novel distance bisection task. 
Subjects are shown the same motion as in the leave-me-alone task. 
However only two agents are displayed. 
One agent is oriented towards chasee and the other one is oriented perpendicular to it. 
Subjects are asked to move a green circle so that it remains on the shortest line between the two agents approximately equidistant to both. 
We expect that the locations chosen by subjects will be displaced in direction away from the wolf. 
The use of multiple tasks with different task demands allows us to identify the perceived displacement more reliably. 
Subjects are not oblivious to the displacement in their explicit judgments and can compensate for the discrepancies. %ms citation needed
The distance bisection task avoids these pitfalls.
The task requires continuous real-time adaption to the changing motion pattern of the two agents. 
There is little opportunity to monitor, reflect and correct the judgment. 
Furthermore, one may object that an observed displacement in location recall task may be due to post-perceptual shift in the memory and is not necessarily relevant to the leave-me-alone task. 
The distance bisection task is not susceptible to this objection.\\
On the other hand, the connection between the distance bisection and leave-me-alone task may not be obvious. 
We suggest that subjects estimate the distance to surrounding agents and then based on this information decide where to move next. 
However, the are many more agents displayed in the leave-me-alone task and the relevant ones are mostly much closer than the average agent in the bisection task. 
In this respect, the location recall judgments are more direct evidence that displacement is also perceived during the leave-me-alone task. \\
So far we framed our study as an attempt to investigate influence of a potential confound in previously published study. 
Such confound would seriously undermine the conclusion of this study, notably the claim that agent's orientation is important cue to detection of chasing and to the perception of goal-directed motion more generally. %ms hier kommen noch ein Paar Referenzen..
However, the current study is also of interest to researchers studying displacement in recall and its relation to action. 
If our claims are correct, this would provide another demonstration of a task where the perceived displacement is relevant to the action-based performance in the task. 
Furthermore we can compare the size of the effects across tasks to see how the size of the reported displacement varies with different task demands.
The outline of this report is as follows. %ms kommt wenn ich das fertig habe

@@

\section{General Methods}

\subsection{Stimulus}
The Stimulus programming followed the description in \citet{gao10} and the examples provided on Brian Scholls web-page. 
The display was divided into four quadrants, each forming a square of size 5.9 degrees. 
Three white objects (1.9 degrees in diameter) moved inside of each quadrant. 
These objects were either bugs (Experiment 1) or darts (Experiment 2). 
A bug was a circle that had two red eyes drawn on top of it. 
(These were two red circles of size 0.19 degrees, located 0.71 degrees from the center of the white circle and 0.49 degress apart.) 
The dart was a white polygon with four vertices. 
Three of them were located at the perimeter and the remaining vertex was at the origin of the circle. 
The wings of the dart formed an angle of 120 degrees.
The subject steered a a green circle (chasee, 1.2 diameter) with computer mouse. 
The movement was confined to lie within a circular area with 11.75 degrees radius. 
The boundary of the circular area was shown on the screen as a thin gray line. 
The motion of the white objects was generated as follows. 
The objects moved at constant speed of 7.8 degrees per second. 
Each object changed its direction at random intervals with 3 direction changes per second on average. 
The new direction was chosen randomly from a range of -45 to 45 degrees around its current direction. 
The objects were pervasive to each other and upon touching the wall of the designated quadrant they bounced off. \\
The orientation was manipulated as follows. 
Agents were oriented either with their eyes/nose directed towards the agent's location or perpendicular to the agent. 
There were two quadrants with perpendicular orientation and two quandrants with head-on oriented agents (wolfpack). 
The type of the quadrant was chosen randomly on each trial. \\

\subsection{Task and Design}
We tested the bug and dart stimuli in two separate session with different subject samples. 
We refer to these as Experiment 1 and Experiment 2 respectively.
An experiment session consisted of 3 blocks with 42 trials per block.\\
Block 1 went as follows. 
At the start of each trial the circular boundary and the initial position of the green circle was displayed. 
Subject initiated the trial by bringing a cross-hair mouse cursor to the green circle and by clicking on its surface. 
Furthermore, we ensured that the agent's initial position was at least 4 degrees away from the nearest agent. 
Then the agents appeared and the subject tried to avoid contact by moving the green circles within the circular boundary. 
This lasted 17 seconds. 
We appended a location judgment task at the end of the trial. 
After the movement ended the agents and chasee remained stationary. 
One agent disappeared. 
The subject then clicked the last location where he saw the missing agent with a cross-hair cursor. 
The missing agent was chosen randomly from the three agents nearest to chasee. 
Whenever possible we tried to select an agent that did not overlap and so that perpendicular and wolfpack agents were chosen approximately equally often within each block. \\
In the second block the rotation point was shifted along anteroposterior axis. The magnitude of this shift was varied across trials. We used trial with displacement of -0.15, -0.05, 0.05, 0.1, 0.15 and 0.2. This range was suggested by pilot studies. Later we adapted the range and used 0.05, 0.1,0.15, 0.2 0.25 and 0.3. Otherwise the trials were identical: The displacement was either perpendicular or towards the green circle. The circles were organized into four quadrants based on the type of the displacement. Finally, each trial was followed by a location judgment task. This allowed us to check whether the displacement manipulation was succesful. In addition to the location judgments in the first two blocks we measured the time spent by the green circle in the quadrants with respect to the orientation of the circles.\\
In third block only two agents from two different quadrants were shown: one with perpendicular and one with head-on orientation. The initial position of green circle set at the mid-point on the line connecting the two white circles. The trial started after the subjects clicked the green circle. The subjects were asked to move the green circle such that they stay at the mid-point between the two agents. In the first 18 trials the orientation was indicated by eyes. In the remaining 24 trials the eyes were not displayed and instead the agents were physically displaced in the direction of agent's orientation. The magnitude of displacement was varied with manipulated values identical to those used in block 2.\\
The third block concluded with 13 trials. A white circle was shown for random interval of 2-3 seconds. Then the crosshair appeared and the the circle either disappeared (in the first five trials) or it remained visible. The subjects were asked to select the position of the white circle with cross hair. On first ten trials the eyes were displayed and their orientation was chosen randomly. \\

\subsection{Eyetracking}
Immediately before taking part in the current experiment the subjects participated as an adult control group in an infant eye-tracking study. 
This experiment took 5-10 minutes. 
Since the subjects were already seated at a calibrated eyetracker we decided to include eyetracking measurements although no analyses of the eyetracking data were planned and none were performed.

\subsection{Procedure}
In block 1 and 2 subjects were instructed to avoid collisions with white agents by moving the green circle with mouse. 
In block 3 subjects were asked to move the green circle such that it stays between the two white agents. \\
The subjects were seated 50-70 cm away from screen (all specifications in degrees of visual angle are referenced to 50 cm distance) of a Tobii T60 display with built-in remote eye-tracker. 
Upon the conclusion of the experiment the subjects were debriefed and dismissed. 
Each block lasted 15 minutes and there was a brief break between them.\\
The experiment was presented and controlled with PsychoPy 1.77 \citep{peirce07} and Tobii SDK 3.0. 
The analyses were performed with Python 2.7 and STAN 2.0.1.

\subsection{Materials and Data}
Subjects gave informed consent to participate in the study and were further given an option to make their data available for sharing with other researchers. 
All subjects agreed. 
The materials and data are available from \url{http://github.com/simkovic/wolfpackRevisited}. 
We  documented all analyses we made in the form of IPython Notebooks which are also available from the online repository.

\subsection{Analyses}
In block 1 and 2, we are interested in the direction of the displacement of the last location judgments of the missing circle with respect to its actual position. We expect four predictors to influence the displacement. First, representational gravity pulls objects down along the vertical axis. Second, under the influence of representational momentum subjects interpolate future positions along the motion trajectory of object and recall these as the last position. Third, we call the displacement in the direction of the circles orientation (as indicated by the eyes' location or physical displacement in the third block), orientation displacement. Finally, we expect that the last position of the green circle (controled by mouse) will influence the judgment. In particular we expect the judgment to be pulled towards the last mouse position. We call this effect the lazy-hand gravity.\\
More formally, let $\mathbf \theta_i =[\theta_{ix}, \theta_{iy}]$ be the vector of the difference between the indicated and the actual position of the vanished circle at trial $i$ along the two axes. Let $\phi_{ik}$ be the angle given by the direction of the predictor at trial $i$, with $k$ the predictor (as listed in the previous paragraph). Then we are interested in the estimates of the regression coefficients $\alpha^1_k$ given by the equation\\

\begin{equation}
\mathbf \theta_i \sim  \mathcal{N}\left(\sum_k \alpha_k \begin{bmatrix} \cos \phi_{ik} \\ \sin \phi_{ik} \end{bmatrix}, I\sigma_\theta^{-2} \right)
\end{equation}

As in \citet{gao10} we compute the proportion of the time spent in the wolfpack quadrants on each trial. We then estimate average proportion for each subject. \\
In block 3 we estimate the displacement as follows. We rotate subjects' judgments in each around the nominal mid-point such that the wolf is located at 180 degrees and perpendicular agent is at 0 degrees. On each trial we discard first 2 seconds and compute the mean displacement during the remaining 15 seconds. We then use this data to estimate the displacement across trials and across subjects.\\
Our general strategy is to first estimate parameters for each subject individually and then to pool estimates with a hierarchical model whenever meaningful. Our within-subject design does not allow us to collect more data points so that in some cases the individual estimates are as robust as we would desire. Hierarchical modeling allows us to pool information across subjects and to obtain more reliable estimates of the individual parameters. More information on hierarchical modeling can be found in \citet{gelman07} and \citet{lee11}.\\

>>>>

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
\usepackage{hyperref}
%% END MACROS SECTION

@@

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Perceived Displacement explains Wolfpack Effect}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matus Simkovic $^{1\ast}$, 
Birgit Tr\"auble $^{1}$
\\
\bf{1}  Universit\"at Heidelberg, Germany
\\
$\ast$ E-mail: matus.simkovic@psychologie.uni-heidelberg.de
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

@@

Recently, Gao and Scholl presented a set of studies \cite{gao09,gao10,gao11} demonstrating that certain aspects of motion influence the detection of chasing and the performance in interactive tasks where subjects control chasee with computer mouse. 
In \cite{gao10} the authors focused on the contribution of agent's orientation. 

@@

For example, in Experiment 2 the subjects controlled a green disc with mouse and tried to avoid contact with a white circle which chased the green circle. 

@@

As a manipulation the authors added seven white darts that were, in half of the trials, oriented towards the green circle and in other half perpendicular to it. 

@@

They compared the two conditions in terms of the proportion of trials where chaser caught the green circle. 

@@

In their Experiment 4 the chaser was the only circle in the display, so it was easily identified. 
However the subjects were additionaly required to avoid contact with darts. 
Escape rate was lower in trials where the darts were oriented towards green circle in comparison to trials where the orientation was perpendicular. 

@@

The authors called the consistent distressing influence of head-on orientation on subject's performance, the wolfpack effect.\\

@@

This is illustrated in figure \ref{fig:rot} on the example of darts used by \cite{gao10}. 
The authors designated the concave vertex (green dot) as the nominal point for rotation. 
Intuitively however, the center of mass (cross) looks like a much better candidate. 
What happens if the perceived center of rotation is displaced in the direction of center of mass?
In this case the rotation around the nominal center is perceived as motion. 
If the agent orients towards chasee this is perceived as a motion towards chasee. 
Consequently, we can contrive an alternative explanation for the results in \cite{gao10}. 
In Experiment 3a the subjects avoided areas with darts oriented towards the prey because they perceived these agents nearer to the green circle. 
In Experiment 4 we assume that subjects estimated the critical distance to the surrounding agents. 
Then based on this estimate they decided where to move the green circle next. 
If the agents' perceived location was shifted towards chasee in wolfpack trials, the subjects would often prematurely leave the current location in exchange for another alternative location. 
This alternative location however would be a bad choice from the point of view of the dart's nominal center (green dot) whose locatio was used to decide whether the green circle was caught or not. 

@@

Still, since the the intended orientation change is perceived as motion, the results can be explained by the influence of motion cues on subjects judgment. 
If properly manipulated the rotation doesn't need to have any influence at all.\\
Experiment 3b in \cite{gao10} was the only experiment that did not use darts. 
The design and results were similar to that of Experiment 3a, except that instead of darts the subject tried to avoid white circles whose orientation was determined by two red dots ('eyes', see e.g. figure \ref{fig:vec}).

@@

If we use the shape of the agent to determine the center of mass, this is identical to the point around which the circle was rotated in the Experiment. 

@@

The literature on memory displacement is relevant here \cite{hubbard05}. 

@@

Crucially, some studies (e.g.\cite{freyd92}) demonstrated that displacement can also be influenced by information about stimulus animacy. 
The memory displacement is studied by asking subjects to report the last position of an object that just disappeared.
This is usually done by asking subjects to select this position with mouse or alternatively by selecting which of several probes is nearest to the location where the object disappeared.
How are such displacements in explicit recall relevant to the implicit interactive tasks used in \cite{gao10}?
According to \cite{hubbard05} (p. 844) ``displacement occurs because it aids in the spatial localization of physical objects and facilitates rapid motor responding to objects in the environment''. 
He further adds (cf.) that ``accurate spatial localization is important for calibrating an observer's response to a stimulus so that a maximally effective and adaptive interaction with that stimulus might be achieved''. 

@@

We can apply this idea to stimuli and tasks in \cite{gao10}. Since agent's head/nose orientation usually predicts the subsequent motion, orientation may be used to predict future position. 

@@

Hence we would predict a memory displacement which is influenced by factors that provide cue to the future motion direction. 

@@

Due to the perceived displacement the subject then avoids the wolfpack areas because the wolfpack stimuli are perceived as closing in on the green circle. 
For similar reasons, we would expect the perceived location of the darts to be shifted even farther towards the nose beyond the position of the center of mass.\\
In the current study we investigate the influence of perceived displacement on subject's performance in the interactive tasks from \cite{gao10}. 
In particular we chose to revisit their task from Experiment 3, the so-called Leave-Me-Alone task in which subjects avoided contact with randomly moving objects in a display separated into areas differing in terms of agents' orientation. 
We chose this Experiment since it was the only one that demonstrated the wolfpack effect with bug stimuli. We use both bugs and darts in separate experiment since these different stimuli provide different cues to displacement.\\

@@

  \item This displacement influences subjects' performance in the Leave-Me-Alone task. 

@@

To test the second claim we shift the rotation center along the anteroposterior axis of the agent and measure how this manipulation influences subject's avoidance of wolfpack areas. We however need to separate the constant influence of the orientation cues from the manipulated displacement. We can use simple circles with no eyes to avoid constant influence. With darts there is no straightforward way to neutralize the constant influence of the shape on the perceived displacement. Instead we use linear regression model to separate the two factors in the analysis.\\
To test the first claim we utilize two tasks. 
First, we append a location recall task to the leave-me-alone task. 
Immediately after each trial of the leave-me-alone task one dart/bug disappears while all the remaining objects freezee at their last position. 
The subject is asked to select the position where the agent disappeared.
Second, we devised a novel distance bisection task. 
Subjects are shown the same motion as in the leave-me-alone task. 
However only two agents are displayed. 
One agent is oriented towards chasee and the other one is oriented perpendicular to it. 
Subjects are asked to move a green circle so that it remains on the shortest line between the two agents approximately equidistant to both. 
We expect that the locations chosen by subjects will be displaced in direction away from the wolf. 
The use of multiple tasks with different task demands allows us to identify the perceived displacement more reliably. 
Subjects are not oblivious to the displacement in their explicit judgments and can compensate for the discrepancies. %ms citation needed
The distance bisection task avoids these pitfalls.
The task requires continuous real-time adaption to the changing motion pattern of the two agents. 
There is little opportunity to monitor, reflect and correct the judgment. 
Furthermore, one may object that an observed displacement in location recall task may be due to post-perceptual shift in the memory and is not necessarily relevant to the leave-me-alone task. 
The distance bisection task is not susceptible to this objection.\\
On the other hand, the connection between the distance bisection and leave-me-alone task may not be obvious. 
We suggest that subjects estimate the distance to surrounding agents and then based on this information decide where to move next. 
However, the are many more agents displayed in the leave-me-alone task and the relevant ones are mostly much closer than the average agent in the bisection task. 
In this respect, the location recall judgments are more direct evidence that displacement is also perceived during the leave-me-alone task. \\
So far we framed our study as an attempt to investigate influence of a potential confound in previously published study. 
Such confound would seriously undermine the conclusion of this study, notably the claim that agent's orientation is important cue to detection of chasing and to the perception of goal-directed motion more generally. %ms hier kommen noch ein Paar Referenzen..
However, the current study is also of interest to researchers studying displacement in recall and its relation to action. 
If our claims are correct, this would provide another demonstration of a task where the perceived displacement is relevant to the action-based performance in the task. 
Furthermore we can compare the size of the effects across tasks to see how the size of the reported displacement varies with different task demands.
The outline of this report is as follows. %ms kommt wenn ich das fertig habe

@@

\section*{General Methods}

\subsection*{Stimulus}
The Stimulus programming followed the description in \cite{gao10} and the examples provided on Brian Scholls web-page. 
The display was divided into four quadrants, each forming a square of size 5.9 degrees. 
Three white objects (1.9 degrees in diameter) moved inside of each quadrant. 
These objects were either bugs (Experiment 1) or darts (Experiment 2). 
A bug was a circle that had two red eyes drawn on top of it. 
(These were two red circles of size 0.19 degrees, located 0.71 degrees from the center of the white circle and 0.49 degress apart.) 
The dart was a white polygon with four vertices. 
Three of them were located at the perimeter and the remaining vertex was at the origin of the circle. 
The wings of the dart formed an angle of 120 degrees.
The subject steered a a green circle (chasee, 1.2 diameter) with computer mouse. 
The movement was confined to lie within a circular area with 11.75 degrees radius. 
The boundary of the circular area was shown on the screen as a thin gray line. 
The motion of the white objects was generated as follows. 
The objects moved at constant speed of 7.8 degrees per second. 
Each object changed its direction at random intervals with 3 direction changes per second on average. 
The new direction was chosen randomly from a range of -45 to 45 degrees around its current direction. 
The objects were pervasive to each other and upon touching the wall of the designated quadrant they bounced off. \\
The orientation was manipulated as follows. 
Agents were oriented either with their eyes/nose directed towards the agent's location or perpendicular to the agent. 
There were two quadrants with perpendicular orientation and two quandrants with head-on oriented agents (wolfpack). 
The type of the quadrant was chosen randomly on each trial. \\

\subsection*{Task and Design}
We tested the bug and dart stimuli in two separate sessions with different subject samples. 
We refer to these as Experiment 1 and Experiment 2 respectively.
An experiment session consisted of 3 blocks with 42 trials per block. 
Overview of all conditions and the amount of obtained data is shown in table \ref{tab.taskoverview}. 
We now provide detailed description. \\
First block went as follows. 
At the start of each trial the circular boundary and the initial position of the green circle was displayed. 
Subject initiated the trial by bringing a cross-hair mouse cursor to the green circle and by clicking on its surface. 
Furthermore, we ensured that the agent's initial position was at least 4 degrees away from the nearest agent. 
Then the agents appeared and the subject tried to avoid contact by moving the green circles within the circular boundary. 
This lasted 17 seconds. 
We appended a location judgment task at the end of the trial. 
After the movement ended the agents and chasee remained stationary. 
One agent disappeared. 
The subject then clicked the last location where he saw the missing agent with a cross-hair cursor. 
The missing agent was chosen randomly from the three agents nearest to chasee. 
Whenever possible we tried to select an agent that did not overlap and so that perpendicular and wolfpack agents were chosen approximately equally often within each block. \\
In the second block the rotation point was shifted along anteroposterior axis. 
The magnitude of this shift was varied across trials. 
Furthermore, we calibrated the manipulation between subjects. 
We wanted to obtain measurement of manipulations that were similar in magnitude to the displacement reported in the recall task and in the distance bisection task. 
In experiment 1 we started with wide displacement range of $(-0.15, -0.05, 0.05, 0.1, 0.15, 0.2)$. 
Later we adapted the range and to $(0.05, 0.1,0.15, 0.2 0.25,0.3)$. 
In experiment 2 we first used values $(0.4, 0.2)$ and later $(-0.4, -0.2)$. 
The reasons for using negative displacement are explained in section \ref{sec.b5}. 
In experiment 1, we adjested the stimuli.
These were now simple white circles without eyes.
Otherwise the trials in block 2 were identical to block 1. 
The displacement was either perpendicular or towards the green circle. 
The agents were organized into four quadrants based on the type of the displacement. Finally, each trial was followed by a location recall task. 
We included the recall task in order to make the first two blocks as similar as possible. 
Furthermore it allowed us to check whether the displacement manipulation was succesful.
In addition to the location judgments in the first two blocks we measured the time spent by the green circle in the quadrants with respect to the orientation of the circles.\\
In third block only two agents from two different quadrants were shown: one with perpendicular and one with head-on orientation. 
The initial position of green circle set at the mid-point on the line connecting the two agents. 
The trial started after the subject clicked on the green circle. 
The subjects were asked to move the green circle such that it stays at the mid-point between the two agents. 
The first 18 trials were analoguous to block 1 in that there was no displacement and in experiment one the circles had eyes. 
In the remaining 24 trials the eyes were not displayed and the agents were physically displaced in the direction of agent's orientation. 
The schedule for the magnitude of displacement was identical to the one used in block 2.\\
A final block concluded with 13 trials. 
An agent was shown for random interval of 2-3 seconds. 
Then the crosshair appeared and the the agent either disappeared (in the first five trials) or it remained visible. 
The subjects were asked to select the position of the agent with cross hair. 
On the first ten trials of experiment 1 the eyes were displayed and their orientation was chosen randomly. 
Similarly the orientation of darts in experiment 2 was chosen randomly.\\

\subsection*{Eyetracking}
Immediately before taking part in the current experiment the subjects participated as an adult control group in an infant eye-tracking study. 
This experiment took 5-10 minutes. 
Since the subjects were already seated at a calibrated eyetracker we decided to include eyetracking measurements although no analyses of the eyetracking data were planned and none were performed.

\subsection*{Procedure}
In block 1 and 2 subjects were instructed to avoid collisions with white agents by moving the green circle with mouse. 
In block 3 subjects were asked to move the green circle such that it stays between the two white agents. \\
The subjects were seated 50-70 cm away from screen (all specifications in degrees of visual angle are referenced to 50 cm distance) of a Tobii T60 display with built-in remote eye-tracker. 
Upon the conclusion of the experiment the subjects were debriefed and dismissed. 
Each block lasted 15 minutes and there was a brief break between them.\\
The experiment was presented and controlled with PsychoPy 1.77 \cite{peirce07} and Tobii SDK 3.0. 
The analyses were performed with Python 2.7 and STAN 2.0.1.

\subsection*{Materials and Data}
Subjects gave informed consent to participate in the study and were further given an option to make their data available for sharing with other researchers. 
All subjects agreed. 
The materials and data are available from \url{http://github.com/simkovic/wolfpackRevisited}. 
We  documented all analyses we made in the form of IPython Notebooks which are also available from the online repository.

\subsection*{Statistical Modeling}
In the reported analyses we follow the approach advocated in \cite{gelman07} and \cite{gelman13}. 
We first design a model captures well the patterns present in the data. 
We use posterior predictive checking to decide whether the model is acceptable. 
We start by fitting a separate model for each subject. 
We then design a hierarchical model that pools data across subjects \cite{lee11}. 
Hierarchical models can be easily formulated within bayesian framework and evaluated with sampling algorithms.
Finally we report and interpret the estimates of population variables that are of theoretical interest. We report the mean estimate and the 95 percent range of the estimate.\\
The main advantage of this approach is that it gives us flexibility to design complex models that match (or at least approach) the complexity of the processes that generated the data.
The model design involves decisions which may considerably alter the posterior estimates and even the conclusions of the study.
We discuss some of the modeling decisions below. 
The analyses are documented and can be replicated with the IPython Notebooks available from the project repository.
In general, virtually all results held at least to some degrees across all the various models we evaluated and as such the main conclusions of this report are robust.

\section*{Results}

\subsection*{Localization and Recall of Static Stimuli}

The data are shown in figure \ref{fig.b5} and summary statistics is given in \ref{tab.b5}. 
As expected, the localization judgments (B, C, E, G) are more precise than recall (A, D, F). Darts show systematic displacement in the direction of the nose (D-G). 
Bugs show no such displacement although the recall task (A) may be worth further investigation. 
Furthermore, the reported origin is displaced further towards the nose than the center of mass (shown as vertical dashed line at 0.13 in figure \ref{fig.b5}). 
Recall that the task with static stimuli was preceded by task with stimuli where the origin was shifted. 
Does the reported location of static stimuli reflect learning and transfer of knowledge from the previous task? 
The second row (D, E) shows data for subjects who saw the origin shifted towards the nose (dots in figure \ref{fig.man}). 
The data are consistent with a learning account. 
However, in third row (F, G) we should expect judgment shifted to the left, but we don't observe any reliable difference between the two groups of subjects. 
Finally, there are notable differences between the subjects. 
The dark purple subject located the origin at the nose and did so also during recall. 
The green subject located the origin behind the dart but only when the stimulus remained on the screen.

\subsection*{Distance Bisection}
\subsubsection*{Bugs}
Figure \ref{fig.b3} shows data from trials where the origin was located at the zero for bug (A) and dart (B) stimuli. 
As expected, most of the measurements and subject medians are located in the lower right quadrant.
We estimated the magnitude of perceived displacement on the x and y axis separately with a hierarchical model (S3.1 with D1.3.1). Subjects judge the wolf as shifted by $0.25 \; [0.11,0.39]$ degrees in direction of its nose (S3.1 $\mu{\mu,x}$). Similar, the orientation of perpendicular agent also influences where subjects put the point of mid-distance. The perpendicular agent is perceived as shifted by $0.09\; [-0.03,0.21]$ \footnote{This would be $-0.1$ in figure \ref{fig.b3}} (S3.1 $\mu{\mu,x}$). The origin displacement on y axis is smaller than on x axis. As can be seen in figure \ref{fig.b3} not all subjects do show this displacement and as a result the estimated value is smaller and noisier.\\
\subsubsection*{Darts}
Bisection judgments for dart stimuli (D2.3.1) are shown in panel B of figure \ref{fig:b3}. As with bugs we fit a hierarchical model (S3.1) for each axis. The same model can be applied to trials where the origin was shifted (D2.3.2 and D2.3.3). The results for x axis are shown in figure \ref{fig:b3reg} A. As expected, the jugments lie on a line with slope approximately 1. We can use this fact to include all bisection data in a a single regression model (S3.2). The fitted regression line is shown in red in figure \ref{fig:b3reg}. The slope is $1.04\; [0.87,1.23]$. Of more interest is the line's offset that gives the perceived displacement independent of the manipulated displacement (and relative to the nominal zero shown in figure \ref{fig:man}). The offset is $0.25\; [0.2,0.29]$. This is considerably larger than zero and larger than the location of center of mass would predict. The data are however consistent with the judgments of the origin of static darts.\\
Similar evaluation was applied for the displacement due to the orientation of the perpendicular dart. The separate estimates for each condition (S3.1) show that the data points do not lie on a line but rather form a sigmoidal curve. If we fit a line observe a slope of $0.88\; [0.63,1.1]$ and a constant offset of $0.15\; [0.09,0.2]$.

\subsubsection*{Circles}
We can fit the regression model (S3.2) to bisection data which used circles with displaced location. In this case there is no orientation cue apart from the manipulated displacement. The slope on the x axis was $0.86\; [0.01,1.74]$ and $0.5 \;[-0.27,1.3]$ for y axis. These estimates are very imprecise due to small sample size. However, we see that also here the displacement on y axis is smaller than on x axis. This effect is thus most probably a property of the task and does not reflect the true displacement.

\subsection*{Location Recall} 
\subsubsection*{Statistical Modeling} 
In block 1 and 2, we are interested in the direction of the displacement of the last location judgments of the missing circle with respect to its actual position. 
We expected four factors to influence the displacement. First, representational gravity pulls objects down along the vertical axis. Second, under the influence of representational momentum subjects interpolate future positions along the motion trajectory of object and recall these as the last position. Third, we expect that the last position of the green circle (controled by mouse) will influence the judgment. In particular we expect the judgment to be pulled towards the last mouse position. We call this effect the lazy-hand gravity. 
Finally, we call the displacement in the direction of the circle's orientation (as indicated by the eyes' location or physical displacement in the third block), orientation displacement.\\
Some of the above mentioned factors are correlated. Therefore we formulate a regression model. 
We assume that the influence of the four factors is linearly additive. This is illustrated in figure \ref{fig:vectors}. 
Several studies of memory displacement demonstrated additivity.
Furthermore we assume bivariate gaussian distribution. Figure \ref{fig:b1} shows that this assumption is not implausible.\\ % citation here
More formally, let $\mathbf h =[h_{x}, h_{y}]$ be the vector of the difference between the indicated and the actual position of the vanished circle along the two axes. Let $\phi_{k}$ be the angle given by the direction of the predictor with $k$ giving the predictor (as listed in the previous paragraph). Then we are interested in the estimates of the regression coefficients $\alpha_k$ given by the equation\\
\begin{equation}
\mathbf h \sim  \mathcal{N}\left(\sum_k \alpha_k \begin{bmatrix} \cos \phi_{k} \\ \sin \phi_{k} \end{bmatrix}, \sigma_h I_{2 \times 2} \right).
\end{equation} 
The regression coefficients are directly interpretable. They give the magnitude of the displacement due to each factor in degrees (for $h$ in degrees).\\
Once more we use hierarchical model to pool the estimates across subjects and to obtain population estimates. A detailed formulation of the model is given in Table \ref{tab:models} (S1.1).
